\chapter{REFERENCES}
\addcontentsline{toc}{chapter}{References}

\begin{thebibliography}{99}

\bibitem{breiman2001}
L. Breiman, ``Random Forests,'' \textit{Machine Learning}, vol. 45, no. 1, pp. 5--32, Oct. 2001.

\bibitem{vaswani2017}
A. Vaswani, N. Shazeer, N. Parmar, J. Uszkoreit, L. Jones, A. N. Gomez, L. Kaiser, and I. Polosukhin, ``Attention Is All You Need,'' in \textit{Advances in Neural Information Processing Systems (NeurIPS)}, vol. 30, 2017.

\bibitem{lee2020}
J. Lee, W. Yoon, S. Kim, D. Kim, S. Kim, C. H. So, and J. Kang, ``BioBERT: A Pre-trained Biomedical Language Representation Model for Biomedical Text Mining,'' \textit{Bioinformatics}, vol. 36, no. 4, pp. 1234--1240, Feb. 2020.

\bibitem{alsentzer2019}
E. Alsentzer, J. R. Murphy, W. Boag, W. Weng, D. Jin, T. Naumann, and M. McDermott, ``Publicly Available Clinical BERT Embeddings,'' in \textit{Proceedings of the 2nd Clinical NLP Workshop}, Minneapolis, MN, 2019.

\bibitem{django2024}
Django Software Foundation, ``Django Documentation, Version 4.2,'' \url{https://docs.djangoproject.com/}, 2024.

\bibitem{react2024}
Meta Open Source, ``React: A JavaScript Library for Building User Interfaces,'' \url{https://reactjs.org/}, 2024.

\bibitem{sklearn2011}
F. Pedregosa, G. Varoquaux, A. Gramfort, V. Michel, B. Thirion, O. Grisel, M. Blondel, P. Prettenhofer, R. Weiss, V. Dubourg, J. Vanderplas, A. Passos, D. Cournapeau, M. Brucher, M. Perrot, and E. Duchesnay, ``Scikit-learn: Machine Learning in Python,'' \textit{Journal of Machine Learning Research}, vol. 12, pp. 2825--2830, 2011.

\bibitem{groq2024}
Groq Inc., ``Groq API Documentation --- Fast AI Inference,'' \url{https://console.groq.com/docs}, 2024.

\bibitem{ieee830}
IEEE, ``IEEE Recommended Practice for Software Requirements Specifications,'' \textit{IEEE Std 830-1998}, 1998.

\bibitem{jwt2015}
M. Jones, J. Bradley, and N. Sakimura, ``JSON Web Token (JWT),'' \textit{RFC 7519}, Internet Engineering Task Force (IETF), May 2015.

\bibitem{pypdf2}
M. Fenniak and contributors, ``PyPDF2: A Python library for PDF file manipulation,'' \url{https://pypdf2.readthedocs.io/}, 2024.

\bibitem{vite2024}
Evan You and contributors, ``Vite: Next Generation Frontend Tooling,'' \url{https://vitejs.dev/}, 2024.

\bibitem{framermotion2024}
Framer Ltd., ``Framer Motion: A production-ready motion library for React,'' \url{https://www.framer.com/motion/}, 2024.

\bibitem{joblib2024}
Joblib Development Team, ``Joblib: Running Python Functions as Pipeline Jobs,'' \url{https://joblib.readthedocs.io/}, 2024.

\end{thebibliography}
